### 2019-07-08 docs/build/latex/ClinicianScheduler.tex
Remove GCal references in manual + rearrange it

%% Generated by Sphinx.
\def\sphinxdocclass{report}
\documentclass[letterpaper,10pt,english]{sphinxmanual}
\ifdefined\pdfpxdimen
   \let\sphinxpxdimen\pdfpxdimen\else\newdimen\sphinxpxdimen
\fi \sphinxpxdimen=.75bp\relax

\PassOptionsToPackage{warn}{textcomp}
\usepackage[utf8]{inputenc}
\ifdefined\DeclareUnicodeCharacter
 \ifdefined\DeclareUnicodeCharacterAsOptional
  \DeclareUnicodeCharacter{"00A0}{\nobreakspace}
  \DeclareUnicodeCharacter{"2500}{\sphinxunichar{2500}}
  \DeclareUnicodeCharacter{"2502}{\sphinxunichar{2502}}
  \DeclareUnicodeCharacter{"2514}{\sphinxunichar{2514}}
  \DeclareUnicodeCharacter{"251C}{\sphinxunichar{251C}}
  \DeclareUnicodeCharacter{"2572}{\textbackslash}
 \else
  \DeclareUnicodeCharacter{00A0}{\nobreakspace}
  \DeclareUnicodeCharacter{2500}{\sphinxunichar{2500}}
  \DeclareUnicodeCharacter{2502}{\sphinxunichar{2502}}
  \DeclareUnicodeCharacter{2514}{\sphinxunichar{2514}}
  \DeclareUnicodeCharacter{251C}{\sphinxunichar{251C}}
  \DeclareUnicodeCharacter{2572}{\textbackslash}
 \fi
\fi
\usepackage{cmap}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amstext}
\usepackage{babel}
\usepackage{times}
\usepackage[Bjarne]{fncychap}
\usepackage{sphinx}

\usepackage{geometry}

% Include hyperref last.
\usepackage{hyperref}
% Fix anchor placement for figures with captions.
\usepackage{hypcap}% it must be loaded after hyperref.
% Set up styles of URL: it should be placed after hyperref.
\urlstyle{same}

\addto\captionsenglish{\renewcommand{\figurename}{Fig.}}
\addto\captionsenglish{\renewcommand{\tablename}{Table}}
\addto\captionsenglish{\renewcommand{\literalblockname}{Listing}}

\addto\captionsenglish{\renewcommand{\literalblockcontinuedname}{continued from previous page}}
\addto\captionsenglish{\renewcommand{\literalblockcontinuesname}{continues on next page}}

\addto\extrasenglish{\def\pageautorefname{page}}

\title{Clinician Scheduler Documentation}
\date{Jul 08, 2019}
\release{1.0}
\author{David Landsman}
\newcommand{\sphinxlogo}{\vbox{}}
\renewcommand{\releasename}{Release}
\makeindex

\begin{document}

\maketitle
\sphinxtableofcontents
\phantomsection\label{\detokenize{manual::doc}}

\begin{sphinxShadowBox}
\sphinxstyletopictitle{Table of Contents}
\begin{itemize}
\item {} 
\phantomsection\label{\detokenize{manual:id5}}{\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{Clinician Configuration}}}
\begin{itemize}
\item {} 
\phantomsection\label{\detokenize{manual:id6}}{\hyperref[\detokenize{manual:creating-a-new-configuration-file}]{\sphinxcrossref{Creating a new configuration file}}}

\item {} 
\phantomsection\label{\detokenize{manual:id7}}{\hyperref[\detokenize{manual:saving-the-configuration-file}]{\sphinxcrossref{Saving the configuration file}}}

\item {} 
\phantomsection\label{\detokenize{manual:id8}}{\hyperref[\detokenize{manual:loading-a-configuration-file}]{\sphinxcrossref{Loading a configuration file}}}

\item {} 
\phantomsection\label{\detokenize{manual:id9}}{\hyperref[\detokenize{manual:adding-a-new-clinician}]{\sphinxcrossref{Adding a new clinician}}}

\item {} 
\phantomsection\label{\detokenize{manual:id10}}{\hyperref[\detokenize{manual:deleting-an-existing-clinician}]{\sphinxcrossref{Deleting an existing clinician}}}

\item {} 
\phantomsection\label{\detokenize{manual:id11}}{\hyperref[\detokenize{manual:editing-an-existing-clinician}]{\sphinxcrossref{Editing an existing clinician}}}
\begin{itemize}
\item {} 
\phantomsection\label{\detokenize{manual:id12}}{\hyperref[\detokenize{manual:scheduling}]{\sphinxcrossref{Scheduling}}}

\end{itemize}

\item {} 
\phantomsection\label{\detokenize{manual:id13}}{\hyperref[\detokenize{manual:generating-a-schedule}]{\sphinxcrossref{Generating a schedule}}}

\item {} 
\phantomsection\label{\detokenize{manual:id14}}{\hyperref[\detokenize{manual:exporting-a-schedule}]{\sphinxcrossref{Exporting a schedule}}}

\item {} 
\phantomsection\label{\detokenize{manual:id15}}{\hyperref[\detokenize{manual:publishing-a-schedule-to-google-calendar}]{\sphinxcrossref{Publishing a schedule to Google Calendar}}}

\item {} 
\phantomsection\label{\detokenize{manual:id16}}{\hyperref[\detokenize{manual:clearing-a-published-schedule}]{\sphinxcrossref{Clearing a published schedule}}}

\end{itemize}

\item {} 
\phantomsection\label{\detokenize{manual:id17}}{\hyperref[\detokenize{manual:sample-output}]{\sphinxcrossref{Sample Output}}}

\end{itemize}
\end{sphinxShadowBox}

\chapter{Clinician Configuration}
\label{\detokenize{manual:clinician-configuration}}\label{\detokenize{manual:id1}}
Before we can generate a schedule, we need to create a configuration file
that specifies which clinicians are available, and how many weeks each
clinician should fulfill.

\section{Creating a new configuration file}
\label{\detokenize{manual:creating-a-new-configuration-file}}
By default, you will get a blank configuration file when you launch
the program. If you would like to discard the changes you have made and
start a configuration file from scratch, simply click on \sphinxtitleref{New Config}.

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/configuration/new\_config.png}{\sphinxincludegraphics{{new_config}.png}}\end{figure}

\begin{sphinxadmonition}{warning}{Warning:}
Unsaved changes to a configuration file will be discarded upon clicking
\sphinxtitleref{New Config}.
\end{sphinxadmonition}

\section{Saving the configuration file}
\label{\detokenize{manual:saving-the-configuration-file}}
When you are ready to save the configuration you created, click on
\sphinxtitleref{Save Config} and choose a place to save your file. Make note of the name
and directory of the file so you could load it in future runs.

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/configuration/save\_config.png}{\sphinxincludegraphics{{save_config}.png}}\end{figure}

\section{Loading a configuration file}
\label{\detokenize{manual:loading-a-configuration-file}}
If you would like to open a previously created configuration file, simply
click on \sphinxtitleref{Open Config}, navigate to the location of the configuration file,
and select it.

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/configuration/load\_config.png}{\sphinxincludegraphics{{load_config}.png}}\end{figure}

\section{Adding a new clinician}
\label{\detokenize{manual:adding-a-new-clinician}}\label{\detokenize{manual:id2}}
1. From the configuration tab, click \sphinxtitleref{New Clinician}. You should see a
form for supplying details.

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/configuration/add\_clinician/step1\_new\_clinician.png}{\sphinxincludegraphics{{step1_new_clinician}.png}}\end{figure}
\begin{enumerate}
\setcounter{enumi}{1}
\item {} 
Fill out the name, email (optional), and divisions that the clinician
will be covering. To add a division you can click on \sphinxtitleref{Add} and a new row
will be added to the table which you can fill out. You can set the minimum
and maximum number of blocks that a clinician can work in a given division.

\end{enumerate}

\begin{sphinxadmonition}{note}{Note:}
A single block corresponds to two weeks.
\end{sphinxadmonition}

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/configuration/add\_clinician/step2\_add\_division.png}{\sphinxincludegraphics{{step2_add_division}.png}}\end{figure}
\begin{enumerate}
\setcounter{enumi}{2}
\item {} 
To delete a row from the table, select the row and then click \sphinxtitleref{Remove}.

\end{enumerate}

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/configuration/add\_clinician/step3\_remove\_division.png}{\sphinxincludegraphics{{step3_remove_division}.png}}\end{figure}
\begin{enumerate}
\setcounter{enumi}{3}
\item {} 
When you are finished entering the data for the clinician, click \sphinxtitleref{Ok}.
You should now see a new entry in the main table for that clinician.

\end{enumerate}

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/configuration/add\_clinician/step4\_add\_clinician.png}{\sphinxincludegraphics{{step4_add_clinician}.png}}\end{figure}

\section{Deleting an existing clinician}
\label{\detokenize{manual:deleting-an-existing-clinician}}
To delete an existing clinician, simply select a row corresponding
to the clinician in the table and click on \sphinxtitleref{Delete Clinician}.

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/configuration/delete\_clinician.png}{\sphinxincludegraphics{{delete_clinician}.png}}\end{figure}

\section{Editing an existing clinician}
\label{\detokenize{manual:editing-an-existing-clinician}}
To edit the information of a clinician, select a row corresponding
to the clinician in the table and click on \sphinxtitleref{Edit Clinician}. You should
see a dialog window where you can change the information. For more
details on how to enter data in the edit dialog, see {\hyperref[\detokenize{manual:adding-a-new-clinician}]{\sphinxcrossref{\DUrole{std,std-ref}{Adding a new clinician}}}}.

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/configuration/edit\_clinician.png}{\sphinxincludegraphics{{edit_clinician}.png}}\end{figure}

\subsection{Scheduling}
\label{\detokenize{manual:scheduling}}

\section{Generating a schedule}
\label{\detokenize{manual:generating-a-schedule}}\label{\detokenize{manual:id3}}
Once you have created a configuration file, you can switch over to the
\sphinxtitleref{Scheduler} tab of the application in order to generate a schedule.
\begin{enumerate}
\item {} 
From the scheduling tab, click on \sphinxtitleref{Load} and select the configuration
file that you generated in {\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}}

\end{enumerate}

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/scheduling/generate\_schedule/step1\_load\_config.png}{\sphinxincludegraphics{{step1_load_config}.png}}\end{figure}
\begin{enumerate}
\setcounter{enumi}{1}
\item {} 
Enter the calendar year for the schedule that you want to generate.

\end{enumerate}

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/scheduling/generate\_schedule/step2\_calendar\_year.png}{\sphinxincludegraphics{{step2_calendar_year}.png}}\end{figure}
\begin{enumerate}
\setcounter{enumi}{2}
\item {} 
\sphinxstylestrong{(Optional)} In order to retrieve the time-off requests that were populated
in Google calendar in \DUrole{xref,std,std-ref}{adding-clinician-requests}, we need to specify the
calendar ID. Open the calendar in your browser, and navigate to the
\sphinxtitleref{Settings and sharing} page.
On the setting page, navigate to the \sphinxtitleref{Integrate calendar} section, and copy
the value under \sphinxtitleref{Calendar ID} to your clipboard.
Paste this value into the \sphinxtitleref{Google Calendar ID} textbox on the \sphinxtitleref{Scheduler}
tab.

\end{enumerate}

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/scheduling/generate\_schedule/step3\_calendar\_id.png}{\sphinxincludegraphics{{step3_calendar_id}.png}}\end{figure}
\begin{enumerate}
\setcounter{enumi}{3}
\item {} 
\sphinxstylestrong{(Optional)} If you supplied a calendar ID in step 3, you can configure
the options \sphinxtitleref{Retrieve Time-off Requests} and \sphinxtitleref{Retrieve Long Weekends} as necessary.
Enabling \sphinxtitleref{Retrieve Time-off Requests} will read the time-off calendar events
from the calendar specified in \sphinxtitleref{Calendar ID}, while enabling
\sphinxtitleref{Retrieve Long Weekends} will read the holiday events from that calendar.

\end{enumerate}

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/scheduling/generate\_schedule/step4\_retrieve\_events.png}{\sphinxincludegraphics{{step4_retrieve_events}.png}}\end{figure}
\begin{enumerate}
\setcounter{enumi}{4}
\item {} 
\sphinxstylestrong{(Optional)} In the case that you only need to generate a schedule for
a subset of the calendar year, you can select the amount of blocks you need
to generate by specifying a value in \sphinxtitleref{Number of Blocks}. By default, the
scheduler will generate a full schedule, corresponding to 26 blocks.

\end{enumerate}

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/scheduling/generate\_schedule/step5\_number\_of\_blocks.png}{\sphinxincludegraphics{{step5_number_of_blocks}.png}}\end{figure}
\begin{enumerate}
\setcounter{enumi}{5}
\item {} 
Click on \sphinxtitleref{Generate}, and after a few moments you should see a preview of
the generated schedule in the table.

\sphinxstylestrong{(Optional)} See \DUrole{xref,std,std-ref}{authentication} in case you encounter problems
with Google calendar at this stage.

\end{enumerate}

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/scheduling/generate\_schedule/step6\_generate\_schedule.png}{\sphinxincludegraphics{{step6_generate_schedule}.png}}\end{figure}

\begin{sphinxadmonition}{note}{Note:}
It is possible that the scheduler will not be able to come up with a
schedule that satisfies your constraints. You can try adjusting
some constraints by changing the minimum and maximum number of blocks of
clinicians in the configuration file. See {\hyperref[\detokenize{manual:clinician-configuration}]{\sphinxcrossref{\DUrole{std,std-ref}{Clinician Configuration}}}}
for more information on changing the configuration file.
\end{sphinxadmonition}

\section{Exporting a schedule}
\label{\detokenize{manual:exporting-a-schedule}}
If you are satisfied with the generated schedule, you can choose to export
it as an Excel file. There are two format options: \sphinxtitleref{Yearly Export} and
\sphinxtitleref{Monthly Export}.

Selecting the \sphinxtitleref{Yearly Export} option will generate an excel file with a single
sheet, displaying the clinicians that are covering a particular division
for a given week or weekend. It is very similar to the table output in
the application itself.

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/scheduling/export\_schedule/export\_yearly.png}{\sphinxincludegraphics{{export_yearly}.png}}\end{figure}

Selecting the \sphinxtitleref{Monthly Export} option will generate a more detailed breakdown
of the schedule, with a separate sheet for every month, detailing which
clinician covers which division on which day.

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/scheduling/export\_schedule/export\_monthly.png}{\sphinxincludegraphics{{export_monthly}.png}}\end{figure}

\section{Publishing a schedule to Google Calendar}
\label{\detokenize{manual:publishing-a-schedule-to-google-calendar}}\label{\detokenize{manual:id4}}
If you are satisfied with the generated schedule, you can publish it
to Google calendar, by clicking on \sphinxtitleref{Publish}.

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/scheduling/publish\_schedule/publish\_schedule.png}{\sphinxincludegraphics{{publish_schedule}.png}}\end{figure}

\begin{sphinxadmonition}{warning}{Warning:}
This process might take some time, especially if the number of events
to be published is quite large.
\end{sphinxadmonition}

\section{Clearing a published schedule}
\label{\detokenize{manual:clearing-a-published-schedule}}
If you want to remove all the published events created by the application,
simply click on \sphinxtitleref{Clear}.

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/scheduling/publish\_schedule/clear\_schedule.png}{\sphinxincludegraphics{{clear_schedule}.png}}\end{figure}

\begin{sphinxadmonition}{warning}{Warning:}
This process might take some time, especially if the number of published
events is quite large.
\end{sphinxadmonition}

\chapter{Sample Output}
\label{\detokenize{manual:sample-output}}
Yearly output format

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/scheduling/sample\_yearly.png}{\sphinxincludegraphics{{sample_yearly}.png}}\end{figure}

Monthly output format

\begin{figure}[H]
\centering
\sphinxhref{\_static/images/scheduling/sample\_monthly.png}{\sphinxincludegraphics{{sample_monthly}.png}}\end{figure}

\renewcommand{\indexname}{Index}
\printindex
\end{document}