### 2019-07-29 paper/tbl/2017_schedule_comparison.tex
Add table for 2017 and comparison table

% Table generated by Excel2LaTeX from sheet '2017'
\begin{table}[h]
	\tiny
	\centering
    \begin{tabular}{c||ccc||ccc}
    	\multicolumn{1}{c||}{\multirow{2}[1]{*}{Week \#}} & \multicolumn{3}{c||}{LP Solution}                                                                                                                                                              & \multicolumn{3}{c}{Historical Data}                                                                                                                                                            \\
    	                                                  &                             HIV                             &                                 ID                                 &                              Weekend                               &                             HIV                             &                                 ID                                 &                              Weekend                               \\ \midrule\midrule
    	                        1                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .608,  .761,  .902}B                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .788,  .788,  .788}D                &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     \\
    	                        2                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .957,  .69,  .518}E                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                        3                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                        4                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .663,  .816,  .557}A                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                        5                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .608,  .761,  .902}B                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
    	                        6                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &            \cellcolor[rgb]{ .663,  .816,  .557}A            &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                        7                         & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
    	                        8                         & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                        9                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .663,  .816,  .557}A                & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     \\
    	                       10                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       11                         &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                       12                         &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .608,  .761,  .902}B                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                       13                         & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                       14                         & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                       15                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
    	                       16                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &              \cellcolor[rgb]{ 1,  .851,  .4}C               &                \cellcolor[rgb]{ .957,  .69,  .518}E                &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     \\
    	                       17                         &            \cellcolor[rgb]{ .788,  .788,  .788}D            & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &              \cellcolor[rgb]{ 1,  .851,  .4}C               &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
    	                       18                         &            \cellcolor[rgb]{ .788,  .788,  .788}D            & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                       19                         &              \cellcolor[rgb]{ 1,  .851,  .4}C               &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                       20                         &              \cellcolor[rgb]{ 1,  .851,  .4}C               &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                       21                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                       22                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                \cellcolor[rgb]{ .957,  .69,  .518}E                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &            \cellcolor[rgb]{ .788,  .788,  .788}D            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                       23                         &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .608,  .761,  .902}B                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
    	                       24                         &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       25                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .663,  .816,  .557}A                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                       26                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                       27                         & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                       28                         & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       29                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .608,  .761,  .902}B                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
    	                       30                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .557,  .663,  .859}F                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                       31                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .663,  .816,  .557}A                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .608,  .761,  .902}B                \\
    	                       32                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &            \cellcolor[rgb]{ .663,  .816,  .557}A            &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
    	                       33                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       34                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .957,  .69,  .518}E                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                       35                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     \\
    	                       36                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &               \cellcolor[rgb]{ .663,  .816,  .557}A                \\
    	                       37                         & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} &                \cellcolor[rgb]{ .957,  .69,  .518}E                &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &            \cellcolor[rgb]{ .663,  .816,  .557}A            &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                       38                         & \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J} &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .663,  .816,  .557}A                &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .608,  .761,  .902}B                &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     \\
    	                       39                         &              \cellcolor[rgb]{ 1,  .851,  .4}C               & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                       40                         &              \cellcolor[rgb]{ 1,  .851,  .4}C               & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .459,  .443,  .443}H                &              \cellcolor[rgb]{ 1,  .851,  .4}C               &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       41                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &              \cellcolor[rgb]{ 1,  .851,  .4}C               &               \cellcolor[rgb]{ .608,  .761,  .902}B                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                       42                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                       43                         &              \cellcolor[rgb]{ 1,  .851,  .4}C               &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
    	                       44                         &              \cellcolor[rgb]{ 1,  .851,  .4}C               &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &            \cellcolor[rgb]{ .663,  .816,  .557}A            &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
    	                       45                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &    \cellcolor[rgb]{ .6,  .2,  1}\textcolor[rgb]{ 1,  1,  1}{J}     &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  \\
    	                       46                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .608,  .761,  .902}B                &            \cellcolor[rgb]{ .608,  .761,  .902}B            & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
    	                       47                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
    	                       48                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &            \cellcolor[rgb]{ .788,  .788,  .788}D            &               \cellcolor[rgb]{ .608,  .761,  .902}B                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    	                       49                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .608,  .761,  .902}B                &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
    	                       50                         &            \cellcolor[rgb]{ .663,  .816,  .557}A            &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .608,  .761,  .902}B                &              \cellcolor[rgb]{ 1,  .851,  .4}C               &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
    	                       51                         &            \cellcolor[rgb]{ .608,  .761,  .902}B            &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &              \cellcolor[rgb]{ 1,  .851,  .4}C               &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E
    \end{tabular}%
	\caption{Comparison of schedules for 2017}
	\label{tbl:2017-schedule-comparison}%
\end{table}%