### 2019-08-19 paper/tbl/runtime_divisions.tex
Change unsatisfied to satisfied requests in table
Add missing tex file for runtime comparison

\begin{table}
	\centering
	\begin{tabular}{|c|c|c|c|c|}
		\hline
		&       \multicolumn{4}{c|}{Number of Divisions}        \\ \hline
		Number of Clinicians &  1   &   2    & 2 (Consecutive Blocks Allowed) &  3   \\ \hline
		10          & 0.22 &  0.51  &              0.26              & 1.49 \\ \hline
		20          & 0.42 & 139.42 &              0.50              &     \\ \hline
		30          & 0.63 &   -    &              0.90              &     \\ \hline
		50          & 1.06 &   -    &              1.30              &     \\ \hline
	\end{tabular}
	\caption{Comparison of program runtime (in seconds) for different number of divisions and total clinicians. ``-'' indicates that a solution was not found after 24 hours.}
	\label{tbl:runtime-divisions-comparison}%
\end{table}