### 2019-08-19 paper/tbl/runtime_divisions.tex
Update division runtime table and rephrase

\begin{table}
	\centering
	\begin{tabular}{|c|c|c|c|c|c|}
		\hline
		                                      & \multicolumn{5}{c|}{Number of Divisions} \\ \hline
		\makecell[l]{Number of \\ Clinicians} &  1   &   2   & 2 (CBA) &  3   & 3 (CBA)  \\ \hline
		                 10                   & 0.18 & 0.94  &  0.24   & 2.02 &   0.30   \\ \hline
		                 20                   & 0.33 & 13.84 &  0.42   &      &   1.17   \\ \hline
		                 30                   & 0.51 &   -   &  0.68   &      &   0.93   \\ \hline
		                 50                   & 0.83 &   -   &  1.12   &      &   1.55   \\ \hline
	\end{tabular}
	\caption{Comparison of program runtime (in seconds) for different number of divisions and total clinicians. ``-'' indicates that a solution was not found after 24 hours. ``CBA'' indicates that \underline{c}onsecutive \underline{b}locks were \underline{a}llowed for these schedules.}
	\label{tbl:runtime-divisions-comparison}%
\end{table}