### 2019-08-19 paper/tbl/runtime_divisions.tex
Work on discussion: split into multiple paragraphs

\begin{table}[htbp]
	\centering
	\begin{tabular}{|c|c||c|c||c|c|}
		\hline
		                                      & \multicolumn{5}{c|}{Number of Divisions} \\ \hline
		\makecell[l]{Number of \\ Clinicians} &  1   &  2   & 2 (CBA) &  3   &  3 (CBA)  \\ \hline
		                 10                   & 0.19 & 0.95 &  0.21   & 1.90 &   0.31    \\ \hline
		                 20                   & 0.31 &  -   &  0.41   &  -   &   0.56    \\ \hline
		                 30                   & 0.57 &  -   &  0.68   &  -   &   0.90    \\ \hline
		                 50                   & 0.78 &  -   &  1.09   &  -   &   1.66    \\ \hline
	\end{tabular}
	\caption{Comparison of program runtime (in seconds) for different number of divisions and total clinicians. `-' indicates that a solution was not found after 24 hours. `CBA' indicates that \underline{c}onsecutive \underline{b}locks were \underline{a}llowed when generating these schedules.}
	\label{tbl:runtime-divisions-comparison}%
\end{table}