### 2019-08-23 paper/tbl/2018_schedule_comparison.tex

% Table generated by Excel2LaTeX from sheet '2018_oldLP'
\begin{table}[htbp]
  \centering
	\begin{adjustbox}{scale=0.8}
    \begin{tabular}{c||ccc||ccc}
    \multicolumn{1}{c||}{\multirow{2}[1]{*}{Week \#}} & \multicolumn{3}{c||}{LP Solution} & \multicolumn{3}{c}{Historical Data} \\
          & HIV   & ID    & Weekend & HIV   & ID    & Weekend \\
    \midrule
    \midrule
    1     & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .459,  .443,  .443}H \\
    2     & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .663,  .816,  .557}A \\
    3     & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .518,  .592,  .69}G \\
    4     & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    5     & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .557,  .663,  .859}F \\
    6     & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ 1,  .851,  .4}C \\
    7     & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .608,  .761,  .902}B \\
    8     & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .518,  .592,  .69}G \\
    9     & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .788,  .788,  .788}D \\
    10    & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .459,  .443,  .443}H \\
    11    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .557,  .663,  .859}F \\
    12    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .663,  .816,  .557}A \\
    13    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .459,  .443,  .443}H \\
    14    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    15    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ 1,  .851,  .4}C \\
    16    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .957,  .69,  .518}E \\
    17    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .788,  .788,  .788}D \\
    18    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .957,  .69,  .518}E \\
    19    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F \\
    20    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ 1,  .851,  .4}C \\
    21    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .663,  .816,  .557}A \\
    22    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ 1,  .851,  .4}C \\
    23    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .788,  .788,  .788}D \\
    24    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ 1,  .851,  .4}C \\
    25    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .518,  .592,  .69}G \\
    26    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .459,  .443,  .443}H & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .788,  .788,  .788}D \\
    27    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .957,  .69,  .518}E \\
    28    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    29    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .788,  .788,  .788}D \\
    30    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .663,  .816,  .557}A \\
    31    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .957,  .69,  .518}E \\
    32    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .557,  .663,  .859}F \\
    33    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    34    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ 1,  .851,  .4}C \\
    35    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .518,  .592,  .69}G \\
    36    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    37    & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .663,  .816,  .557}A \\
    38    & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .957,  .69,  .518}E \\
    39    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .788,  .788,  .788}D \\
    40    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    41    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .518,  .592,  .69}G \\
    42    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .557,  .663,  .859}F \\
    43    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ 1,  .851,  .4}C \\
    44    & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
    45    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .957,  .69,  .518}E & \cellcolor[rgb]{ .957,  .69,  .518}E \\
    46    & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .557,  .663,  .859}F & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .663,  .816,  .557}A \\
    47    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .788,  .788,  .788}D \\
    48    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .518,  .592,  .69}G \\
    49    & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .663,  .816,  .557}A & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ .557,  .663,  .859}F \\
    50    & \cellcolor[rgb]{ .788,  .788,  .788}D & \cellcolor[rgb]{ 1,  .851,  .4}C & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .518,  .592,  .69}G \\
    51    & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .608,  .761,  .902}B & \cellcolor[rgb]{ .518,  .592,  .69}G & \cellcolor[rgb]{ .957,  .69,  .518}E \\
    \end{tabular}%
	\end{adjustbox}
	\caption{Comparison of schedules for 2018}
	\label{tbl:2018-schedule-comparison}%
\end{table}%