### 2019-08-24 paper/paper.abstract.tex
SM edits abstract
Well done!

Scheduling of personnel in a hospital environment is vital to improving the service provided to patients and balancing the workload assigned to clinicians. Many approaches have been tried and successfully applied to generate efficient schedules in such settings. However, due to the computational complexity of the scheduling problem in general, most approaches resort to heuristics to find a non-optimal solution in a reasonable amount of time. We designed an integer linear programming formulation to find an optimal schedule in a clinical division of a hospital. It mitigates computational complexity issues by maintaining a minimal set of constraints, yet still provides the flexibility necessary to adapt the formulation to a variety of clinical divisions. \\

We then conducted a case study for our approach using data from the Infectious Diseases division at St. Michael's Hospital in Toronto, Canada. We analyzed and compared the results of our approach to manually-created schedules at the hospital, and found improved adherence to departmental constraints and clinician preferences. We then used simulated data to examine the the sensitivity of the runtime of our linear program for various parameters and observed reassuring results, signifying the practicality and generazability of our approach in different real-world scenarios.