### 2019-08-29 paper/paper.tex
Add preliminary title + author/affil list

\documentclass[]{article}
\input{paper.includes.tex}
\input{paper.definitions.tex}

\title{A flexible integer linear programming formulation for scheduling clinician on-call service in hospitals}

\author[1, 5]{David Landsman}
\author[5]{Huiting Ma}
\author[5]{Jesse Knight}
\author[6]{Kevin Gough}
\author[2, 3, 4, 5]{Sharmistha Mishra}

\renewcommand\Affilfont{\itshape\small}
\affil[1]{Department of Computer Science, University of Toronto}
\affil[2]{Department of Medicine, Division of Infectious Disease, University of Toronto}
\affil[3]{Institute of Health Policy, Management and Evaluation, Dalla Lana School of Public Health, University of Toronto}
\affil[4]{Institute of Medical Sciences, University of Toronto}
\affil[5]{MAP Centre for Urban Health Solutions, Unity Health Toronto}
\affil[6]{Department of Medicine, St. Michael's Hospital}

\begin{document}
	\maketitle

	\begin{abstract}
		\input{paper.abstract.tex}
	\end{abstract}

	\section{Introduction} \label{sec:introduction}
		\input{paper.intro.tex}
	\section{Problem} \label{sec:problem}
		\input{paper.problem.tex}
	\section{Methods} \label{sec:methods}
		\input{paper.methods.tex}
	\section{Results} \label{sec:results}
		\input{paper.results.tex}
	\section{Discussion} \label{sec:discussion}
		\input{paper.discussion.tex}
	\bibliographystyle{unsrt}
	\bibliography{references}
\end{document}