### 2019-09-04 paper/paper.discussion.tex
Minor typo and other fixes

In this paper, we present a simple, yet flexible, integer linear programming formulation to generate schedules for clinical departments at hospitals.
The challenge in applying ILP to the task of scheduling clinicians lies in the computational complexity of finding an optimal solution. As the size of the scheduling problem grows, due to a larger roster of clinicians or more complicated constraints, the time it takes to generate an optimal schedule may grow exponentially.
%The challenge in applying such an approach to this task lies in the fact that ILP is an NP-hard problem. %specify what 'such an approach' and 'this task' are... the first part of the sentence was a bit hard to follow. Define what you mean by NP-hard problem and cite --> i.e. NP-hard = the following sentence about how time to optimal solution grows exponentially? if yes, then instead of 'As such' - say 'That is,...or "An NP-hard problem means that..."
As a result, the majority of approaches taken to create schedules for similar scenarios tend to use heuristics in order to find an approximately optimal solution in a shorter time \cite{burke_state_2004}. \\  %cite/reference statement (i.e. the 'vast majority'...) I can't remember if the introduction details the heuristics --> but if yes, then great. if no, then include that review of the prior literature here in discussion. I think as you said - the journals we are submitting to seem to have more of those details in the introduction.

%double-check all abbreviations have been defined. I would prefer limiting abbreviations to no more than 3 at most if possible. We are ok for word count in this paper :).

We presented a formulation that includes both hard constraints to ensure the schedule satisfies hospital and logistics requirements, and a multi-goal objective function to satisfy the work preferences of clinicians.
%Our formulation includes hard constraints to ensure the schedule satisfies hospital and logistics requirements. It also aims to satisfy the work preferences of clinicians in a clinical department by optimizing a multi-goal objective function. 
Although we restricted our use of the formulation to a minimal set of constraints for the particular needs of the case study (St. Michael's Hospital Division of Infectious Diseases) - our formulation can be adapted to various clinical departments at different hospitals. The flexibility of our ILP allows changing the number of services provided in a division, the length of a work block, clinicians' preference for block to weekend adjacency as well as clinicians' requests for time off. \\

When comparing the optimal schedule generated by our tool to the manually-created schedules at St. Michael's Hospital, we found that the ILP formulation was always able to find an optimal schedule satisfying all required hard constraints, unlike the manual schedule, which often did not take all constraints into account. Moreover, due to the multi-goal objective function in the ILP, the tool was able to maximize and fulfill the majority of clinician preferences and requests, more so than the manually-created schedule. These observations reinforce the benefits of automated tools when generating schedules in hospital departments to balance the work-load of clinicians and improve the service provided to patients. The use of automated tools alleviates the time spent on designing the schedule by hand, and provides clinical departments with a more fair distribution of work that helps improve the overall satisfaction of both employees and patients \cite{silvestro_evaluation_2000}. \\  % how the finding compares to the wider literature when automated compared to manual? What does it mean re: "so what" - human error in generating tools, etc. or challenges in heuristics used manually by people? include citations this last sentence in particular - what dos it reinforce exactly - what are the benefits?

In our simulations, we also found that increasing the number of requests per clinician does not affect the runtime of the algorithm, highlighting the flexibility of the tool in incorporating clinician preferences. Further, we saw that the algorithm can accommodate an increase in time-horizon up to four years with little impact on runtime, which can be applied to departments that generate schedules far in advance. One key limitation we identified was the sensitivity of the runtime to larger numbers of services offered in a single department. Such cases are unlikely to be encountered in the real-world because most clinical departments tend to provide a single service or at most two services by the same roster of clinicians [{\color{red}SM/Kevin, could you help with reference here?}]. One solution to mitigate the runtime issues created by a larger number of services would be relaxing the constraints that prevent assignment of consecutive blocks, followed by manual readjustment from the generated schedule. %what do we mean by certain constraints? specify the other constraints?
Overall, our sensitivity analyses using simulated data provide reassurance that the ILP formulation can be applied to schedule clinicians across real-world variability between clinical departments. %SM - will look for reference and we can ask Kevin Gough too.
Next steps include expanding the generalizability of the tool beyond smaller clinical departments to larger departments within and outside of health-care - especially those that provide multiple services in parallel for patients and other clients.