### 2019-09-07 paper/paper.intro.tex
Add hard line breaks for easier editing

Hospital departments provide services where patient needs, and thus the system's
demands, often exceed the available supply. In particular, on-call schedules for
a fixed number of health-care providers are central to the efficient running of
hospitals. Carefully allocated schedules are meant to simultaneously ensure
sufficient resources are provided to patients while not overworking clinicians
to prevent costly mistakes. It is common practice for on-call schedules in
hospitals to be created manually, yet manually-created schedules are prone to
errors and potential for various biases. First, when there is a large number of
clinicians in a single department, or the constraints that need to be satisfied
by the department are very complex, a manual method may not provide a schedule
that satisfies all clinicians. Second, such methods are likely to overlook
certain constraints that must be maintained to have an operational department,
such as preventing many consecutive work blocks from being assigned or ensuring
clinicians are allocated a specific amount of work blocks throughout the year.
Third, manual scheduling is often time-consuming for the person developing the
schedule. For these reasons, it is important to develop automated methods that
can generate optimal schedules that satisfy the given constraints of the
hospital department. \\
%For example, it is important that a hospital department allocates its
%resources, such as the availability of a finite number of clinicians,
%optimally, to ensure the best possible service for its patients.

Automated methods to optimize schedules have been studied and applied in many
industries, including transportation \cite{aickelin_improved_2006,
	goel_truck_2012, gunther_combined_2010}, manufacturing
\cite{al-yakoob_mixed-integer_2007, al-yakoob_column_2008,
	alfares_simulation_2007}, retail \cite{chapados_retail_2011,
	nissen_automatic_2010} and military \cite{horn_scheduling_2007,
	laguna_modeling_2005}. Of special interest to a clinician on-call scheduling
problem are the approaches to schedule nurses, who often work in shifts. In the
nurse scheduling problem, the goal is to find an optimal assignment of nurses to
shifts that satisfies all of the hard constraints, such as hospital regulations,
and as many soft constraints as possible, which may include nurse preferences. A
wide variety of approaches, including exact and heuristic approaches, have been
used to solve the nurse scheduling problem: integer linear programming
\cite{azaiez_0-1_2005, trilling_nurse_2006, widyastiti_nurses_2016}, network
flows \cite{el_adoly_new_2018}, genetic algorithms
\cite{aickelin_exploiting_2000, jan_evolutionary_2000, kawanaka_genetic_2001},
simulated annealing \cite{jaszkiewicz_metaheuristic_1997}, and artificial
intelligence \cite{abdennadher_nurse_nodate, li_hybrid_2003}. A comprehensive
literature review of these and other methods applied to nurse rostering is
presented in \cite{burke_state_2004}. \\

%An extensive literature review of these and other methods is presented by [??].
%We will briefly summarize the main ideas of some of these approaches. \\ SM -
%don't need to introduce that you will do this for this type of paper I think -
%would do in a thesis chapter though.

Many of the approaches to nurse scheduling were designed to satisfy the
requirements of a specific hospital department which causes a large number of
variables and constraints to be incorporated into the problem formulation. While
these department-specific approaches allow end-users to find precise schedules
that satisfy the needs of the department and the preferences of the nurses and
clinicians in that department, they are difficult to readily adapt to other
departments in the same hospital or other hospitals. % explain why hard to
%adapt?... what makes their generalizablility/adaptability limited?
Moreover, the large number of variables and constraints also leads to
computational complexity issues \cite{goos_complexity_1996}, especially when
using exact methods for finding the solution. In this paper, we tackle a version
of the nurse scheduling problem arising from a case study of one clinical
division, providing two different services simultaneously (general infectious
disease (ID) consults; and HIV consults service) at St. Michael's Hospital in
Toronto, Canada. Our goal is to (1) present a simple integer linear programming
(ILP) formulation for the scheduling problem as developed for the hospital, and
describe the adaptability of the formulation to solving similar problems in
other departments; (2) compare the performance of the ILP scheduler to the
results of the manual approach; and (3) analyze the robustness of the algorithm
in difficult instances of the problem. \\
% present a simple formulation for the problem developed and tested at the
%hospital after switching from a manual approach to scheduling; and (2) analyze
%the performance of integer linear programming in solving difficult instances of
%the problem and compare the results with those of the manual approach; and (3)
%describe the adaptability of the formulation as a basic framework for solving
%similar problems in other departments. \\ %make #3 an objective

We begin by describing the problem in Section \ref{sec:problem}, and presenting
our ILP formulation in Section \ref{sec:methods}. Next, we compare the results
of our formulation to manually-created schedules, and evaluate the performance
of the algorithm on simulated data in Section \ref{sec:results}. Finally, we
discuss and interpret the results in Section \ref{sec:discussion}. % list the
%main contents/elements of the paper.