### 2019-09-07 paper/tbl/runtime_services_clinicians.tex

\begin{table}[htbp]
	\centering
	\begin{tabular}{|c|c||c|c||c|c|}
		\hline
		                                      &  \multicolumn{5}{c|}{Number of Services}  \\ \hline
		\makecell[l]{Number of \\ Clinicians} &  1   &    2    & 2 (without NCB) &  3   & 3 (without NCB) \\ \hline
		                 10                   & 0.16 &  0.74   &  0.16   & 1.40 &  0.23   \\ \hline
		                 20                   & 0.25 & 7468.86 &  0.32   &  -   &  0.43   \\ \hline
		                 30                   & 0.42 &    -    &  0.49   &  -   &  0.66   \\ \hline
		                 50                   & 0.62 &    -    &  0.82   &  -   &  1.14   \\ \hline
	\end{tabular}
	\caption{Comparison of program runtime (in seconds) for different number of services and total clinicians in the division. `-' indicates that a solution was not found within 24 hours. `without NCB' indicates that the No Consecutive Blocks (NCB) constraint was relaxed.}\label{tbl:runtime-services-clinicians-comparison}%
\end{table}