### 2019-09-07 paper/paper.tex

\documentclass[]{article}
\input{paper.includes.tex}
\input{paper.definitions.tex}

\title{A flexible integer linear programming formulation for scheduling
	clinician on-call service in hospitals}

\author[a, b]{David Landsman}
\author[a]{Huiting Ma}
\author[a]{Jesse Knight}
\author[c]{Kevin Gough}
\author[a, c, d, e]{Sharmistha Mishra}

\renewcommand\Affilfont{\itshape\small}
\affil[a]{MAP Centre for Urban Health Solutions, Unity Health Toronto, Toronto,
	ON, Canada}
\affil[b]{Department of Computer Science, University of Toronto, Toronto, ON,
	Canada}
\affil[c]{Department of Medicine, Division of Infectious Disease, St.\ Michael's
	Hospital, University of Toronto, Toronto, ON, Canada}
\affil[d]{Institute of Health Policy, Management and Evaluation, Dalla Lana
	School of Public Health, University of Toronto, Toronto, ON, Canada}
\affil[e]{Institute of Medical Sciences, University of Toronto, Toronto, ON,
	Canada}

\begin{document}
	\maketitle

	\begin{abstract}
		\input{paper.abstract.tex}
	\end{abstract}

	\section{Introduction}\label{sec:introduction}
	\input{paper.intro.tex}
	\section{Problem}\label{sec:problem}
	\input{paper.problem.tex}
	\section{Methods}\label{sec:methods}
	\input{paper.methods.tex}
	\section{Results}\label{sec:results}
	\input{paper.results.tex}
	\section{Discussion}\label{sec:discussion}
	\input{paper.discussion.tex}
	\section{Acknowledgements}\label{sec:acknowledgements}
	\input{paper.acknowledgements.tex}
	\section{Funding}\label{sec:funding}
	\input{paper.funding.tex}

	\bibliographystyle{unsrt}
	\bibliography{references}
\end{document}