### 2019-09-13 paper/tbl/runtime_services_clinicians.tex
tables & figures cleanup

\begin{table}[htbp]
	\centering
 	\caption{Comparison of program runtime (in seconds) for different numbers of services and total clinicians in the division.}%
  \label{tbl:runtime-services-clinicians-comparison}%
	\begin{tabular}{|c|c||c|c||c|c|}
		\toprule
		                                      &  \multicolumn{5}{c|}{Number of Services}  \\ \midrule
		\makecell[l]{Number of \\ Clinicians} &  1   &    2    & 2 (without NCB) &  3   & 3 (without NCB) \\ \midrule
		                 10                   & 0.16 &  0.74   &  0.16   & 1.40 &  0.23   \\ \hline
		                 20                   & 0.25 & 7468.86 &  0.32   &  --  &  0.43   \\ \hline
		                 30                   & 0.42 &    --   &  0.49   &  --  &  0.66   \\ \hline
		                 50                   & 0.62 &    --   &  0.82   &  --  &  1.14   \\ \bottomrule
	\end{tabular}\\[1em]
  \footnotesize\raggedright
  Notes:
  ``--'' indicates that no solution was found within 24 hours;
  ``without NCB'' indicates that the No Consecutive Blocks (NCB) constraint was relaxed.
\end{table}