### 2019-10-11 paper/paper.problem.tex

At St.\ Michael's Hospital, the division of infectious diseases offers
separate but concurrent services for general ID consultations and for HIV consultations. Each
service provides clinical care throughout the year, during regular working hours
and on weekends and holidays. The schedule is created in advance,
outlining all work-week and weekend shifts for the full year.
In the yearly schedule clinicians are assigned to ``blocks'' of
two consecutive regular work weeks and individual weekends. Apart
from long (holiday) weekends, a work week starts on Monday at 8 A.M. and ends on
Friday at 5 P.M. Accordingly, weekend service starts on Friday at 5 P.M., and
ends on Monday at 8 A.M. During the weekend, ID and HIV consultation services
are combined and provided by one clinician. During the regular work week the ID
and HIV services are led by one clinician each.
Therefore, our objective is to assign a single clinician to cover
each service for each block and each weekend of the year, while 
additionally ensuring a balanced workload.

In most scheduling problems, the constraints can be divided into hard and soft
constraints. Hard constraints must be satisfied by any candidate solution, while
soft constraints can be used to select a more favourable solution from a set of
candidate solutions. Typically, soft constraints are therefore encoded as objective
functions which are
maximized when finding a solution. In the case of the clinician
scheduling problem, we chose departmental regulations and workload balance as hard constraints,
while clinician preference and requests serve as soft constraints.
After the schedule is generated, clinicians may
exchange certain weeks or days throughout the year, to fulfill any missed requests.

The following are the departmental and workload constraints placed on the clinician assignments. 
First, each clinician
must work between a minimum and maximum number of blocks of each service during the year.
For instance, one clinician might have to
provide 3-5 blocks of general ID service and 2-3 blocks of HIV service
throughout the year. These limits may be different for each clinician,
and they may change from year to year as
the number of clinicians in the department changes. Second, the schedule
should not assign a clinician to work for two consecutive blocks or two
consecutive weekends. The schedule should also distribute regular
weekends and holiday weekends each equally among all clinicians.

In addition to balancing the workload among clinicians, the schedule
should accommodate their preferences. Clinicians provide their requests for
time off before schedule generation so that the requests may be integrated
into the schedule. Clinicians may specify days, weeks or weekends off, with the
understanding that any blocks overlapping with their request will be assigned to
a different clinician where possible. For example, if a clinician only requests
a given Monday and Tuesday off, the schedule should avoid assigning the
entire block to that clinician. Clinicians also typically prefer to have their weekend and
block assignments side by side, so the schedule should accommodate this where possible.
A summary of the outlined constraints is given in
Table~\ref{tbl:constraint-summary}.

\begin{table}[h]
	\centering
	\caption{Summary of the constraints for the clinician scheduling problem}%
  \label{tbl:constraint-summary}
	\begin{tabular}{ l c l l }
		\toprule
		\textbf{Constraint Name} & \textbf{Abbreviation} & \textbf{Description}                                                                                                                 & \textbf{Type} \\ \midrule
		Block Coverage                                                                  & BC                    & \makecell[l]{each service needs to have
			exactly \\ one clinician that covers any given block}                                     & Hard          \\ \hline
		Weekend Coverage                                                                & WC                    & \makecell[l]{every weekend needs to have
			exactly \\ one clinician that covers it}                                                 & Hard          \\ \hline
		Min/Max                                                                         & MM                    & \makecell[l]{for a given service, each
			clinician can only \\ work between the minimum and maximum \\ number of allowed
			blocks} & Hard          \\ \hline
		No Consecutive Blocks                                                           & NCB                   & \makecell[l]{any clinician should not work \\
			two consecutive blocks, across all services}                                        & Hard          \\ \hline
		No Consecutive Weekends                                                         & NCW                   & \makecell[l]{any clinician should not work two
			consecutive \\ weekends}                                                           & Hard          \\ \hline
		Equal Weekends                                                                  & EW                    & \makecell[l]{weekends should be equally
			distributed \\ between clinicians}                                                        & Hard          \\ \hline
		Equal Holidays                                                                  & EH                    & \makecell[l]{long weekends should be equally
			distributed \\ between clinicians}                                                   & Hard          \\ \hline
		Block Requests                                                                  & BR                    & \makecell[l]{each clinician can request to be
			off service \\ during certain blocks throughout the year}                           & Soft          \\ \hline
		Weekend Requests                                                                & WR                    & \makecell[l]{each clinician can request to be
			off service \\ during certain weekends throughout the year}                         & Soft          \\ \hline
		Block-Weekend Adjacency                                                         & BWA                   & \makecell[l]{the block and weekend
			assignments of a given \\ clinician should be adjacent}                                        & Soft          \\ \bottomrule
	\end{tabular}
\footnotesize\raggedright
Hard constraints must be satisfied by any candidate schedule. 
Soft constraints are optionally satisfied, and are used to rank the set of candidate solutions.
\end{table}