### 2019-10-11 paper/paper.tex
Add corresponding author + keywords

\documentclass[]{article}
\input{paper.includes.tex}

\title{A flexible integer linear programming formulation for scheduling
	clinician on-call service in hospitals}

\author[a, b]{David Landsman}
\author[a]{Huiting Ma\thanks{Contributed equally to this work.}}
\author[a, c]{Jesse Knight\printfnsymbol{1}}
\author[d]{Kevin Gough}
\author[a, c, d, e]{Sharmistha Mishra\thanks{Corresponding author (\texttt{sharmistha.mishra@utoronto.ca})}}

\renewcommand\Affilfont{\itshape\small}
\affil[a]{MAP Centre for Urban Health Solutions, St.\ Michael's Hospital, 
	Unity Health Toronto, Toronto, ON, Canada}
\affil[b]{Department of Computer Science, University of Toronto, Toronto, ON,
	Canada}
\affil[c]{Institute of Medical Sciences, University of Toronto, Toronto, ON,
	Canada}
\affil[d]{Department of Medicine, Division of Infectious Disease, St.\ Michael's
	Hospital, Unity Health Toronto, Toronto, ON, Canada}
\affil[e]{Institute of Health Policy, Management and Evaluation, Dalla Lana
	School of Public Health, University of Toronto, Toronto, ON, Canada}

\begin{document}
	\maketitle

	\begin{abstract}
		\input{paper.abstract.tex}
	\end{abstract}
	\keywords{Clinician scheduling, Integer linear programming, Exact method, Optimization, Clinician requests}

	\section{Introduction}\label{sec:introduction}
	\input{paper.intro.tex}
	\section{Problem}\label{sec:problem}
	\input{paper.problem.tex}
	\section{Methods}\label{sec:methods}
	\input{paper.methods.tex}
	\section{Experiments}\label{sec:experiments}
	\input{paper.experiments.tex}
	\section{Discussion}\label{sec:discussion}
	\input{paper.discussion.tex}
	\section{Acknowledgements}\label{sec:acknowledgements}
	\input{paper.acknowledgements.tex}
	\section{Funding}\label{sec:funding}
	\input{paper.funding.tex}

	\bibliographystyle{unsrt}
	\bibliography{references}
\end{document}