### 2019-10-17 paper/blinded/paper.blinded.tex
Add blinded version of paper

\documentclass[]{article}
\input{paper.includes.blinded.tex}

\title{A flexible integer linear programming formulation for scheduling
	clinician on-call service in hospitals}

\begin{document}
	\maketitle
	\begin{abstract}
		\input{paper.abstract.blinded.tex}
	\end{abstract}
	\keywords{Clinician scheduling, Integer linear programming, Exact method, Optimization, Clinician requests}

	\section{Introduction}\label{sec:introduction}
	\input{paper.intro.blinded.tex}
	\section{Problem}\label{sec:problem}
	\input{paper.problem.blinded.tex}
	\section{Methods}\label{sec:methods}
	\input{paper.methods.blinded.tex}
	\section{Experiments}\label{sec:experiments}
	\input{paper.experiments.blinded.tex}
	\section{Discussion}\label{sec:discussion}
	\input{paper.discussion.blinded.tex}

	\bibliographystyle{unsrt}
	\bibliography{references.blinded}
\end{document}