### 2019-10-17 paper/blinded/titlepage.tex
Add separate title page

\documentclass[]{article}
\usepackage{authblk}
\usepackage[margin=3cm]{geometry}

\makeatletter
\newcommand{\printfnsymbol}[1]{%
	\textsuperscript{\@fnsymbol{#1}}%
}
\makeatother

\title{A flexible integer linear programming formulation for scheduling
	clinician on-call service in hospitals}
\date{}

\author[a, b]{David Landsman}
\author[a]{Huiting Ma\thanks{Contributed equally to this work.}}
\author[a, c]{Jesse Knight\printfnsymbol{1}}
\author[d]{Kevin Gough}
\author[a, c, d, e]{Sharmistha Mishra\thanks{Corresponding author (\texttt{sharmistha.mishra@utoronto.ca})}}

\renewcommand\Affilfont{\itshape\small}
\affil[a]{MAP Centre for Urban Health Solutions, St.\ Michael's Hospital, 
	Unity Health Toronto, Toronto, ON, Canada}
\affil[b]{Department of Computer Science, University of Toronto, Toronto, ON,
	Canada}
\affil[c]{Institute of Medical Sciences, University of Toronto, Toronto, ON,
	Canada}
\affil[d]{Department of Medicine, Division of Infectious Disease, St.\ Michael's
	Hospital, Unity Health Toronto, Toronto, ON, Canada}
\affil[e]{Institute of Health Policy, Management and Evaluation, Dalla Lana
	School of Public Health, University of Toronto, Toronto, ON, Canada}

\begin{document}
	\maketitle
\end{document}