### 2020-07-06 paper/paper.comparison.tex
Update paper.comparison.tex

Prior work by [authors] have   %avoid editorial type statements without citations or numbers, best to state as objective statements
provided categories to organize the types of problems and approaches
commonly used for real-world nurse rostering.
These categories highlight the similarities  %SM what does "this" refer to? this paper or the prior literature? [always use a concrete noun after "this or these" to help the reader :)
and differences
among computationally easy and hard instances of nurse rostering and 
help end-users determine
which algorithms
are best suited to each type of problem.

In XXXyear, De Causmaecker and Berghe [ref] addressed a gap in the literature 
surrounding a central categorization system for the nurse rostering problems identified in the literature.
The authors categorized
nurse rostering problems using a similar approach used in theoretic scheduling problems,
based around the machine environment, job characteristics and constraints,
and optimization objective. They also created a classification that takes into account
hard and soft constraints related to nurse preferences, consecutive 
assignments, coverage and various flexibilities in terms of shift and
skill types. De Causmaecker and Berghe went on to describe where recent examples of the nurse rostering
literature fit into their classification. %SM: give examples of their recent examples.

% DL - maybe this should be in complexity section? %SM: agree, move to complexity section.
Vanhoucke and Maenhout [ref] analyze and compare nurse rostering instances
based on complexity indicators arising from problem constraints. They consider
problem size (in terms of number of nurses, shifts and days), preference distribution
of the nurses, coverage distribution and time-related constraints (such as weekend
and consecutive constraints) as the four indicators for problem complexity.
Then, they generate sample problems corresponding to all four complexity indicators
and measure the effect of increasing complexity on CPU time required for solving each
problem.

[maybe 1 additional paragraph on categorization systems?] %SM: yes, describe a bit more or a table/box as the list of categories. That way - flows into next paragraph on where to place the SMH solution.
%SM i.e. describe/define each category

We follow the categorization by [ref-De Causmaecker and Berghe] to situate our clinician
scheduling problem. Our problem includes 
personnel environment constraints targeting the availability of clinicians
([list which constraints...]),
sequences of assignments ([...]), and balance of rosters ([...]) 
which fall under the categories $\alpha : A$, $\alpha : S$ and $\alpha : B$,  %are these the categories? think other way around -- these fall into the named categories from De Causeameker, etc.?
respectively. 
For work characteristics, the relevant category for our problem are the range
constraints, $\beta : R$, 
as we restrict the workload of each clinician between a minimum and maximum
allowed blocks ([constraint...]).
Lastly, for the optimization objective our problem corresponds to the category
$\gamma : P$, as our optimization objective
aim to adhere to clinician preferences. Therefore the appropriate categorization
for our problem is $ASB/R/P$.

Using the classification in [ref-De Causmaecker and Berghe], we compare our clinician  %use the term classification or categorization, but not both
scheduling problem and solution approach with existing literature. Azaiez and Al Sharif [ref]
describe a nurse rostering problem to schedule nurses on a per-day basis, for a total period
of 28 days. De Causmaecker and Berghe classify this problem as $ASBN/V2/PL$.
In this problem, they are required to assign a minimum number of staff members
per day, unlike in our clinician scheduling problem. Similar to our problem, they include  %who are 'they'? 
hard constraints to prevent consecutive assignments, and enforce a minimum and maximum 
number of shifts for each nurse. When formulating the optimization goal, Azaiez and Al Sharif
focus on balancing the workload among nurses, preventing day-night consecutive shifts and avoiding
off-on-off assignments. Their formulation does not incorporate nurse preferences
for time off into the model.

In [ref-Bilgin 2008], Bilgin et al address a nurse rostering problem using local search. 
The problem presented is classified as $ASBI/RVNO/PLR$.  %problem presented by whom?
Some of the constraints they incorporate include coverage constraints, overlapping assignment  %who are 'they'?
constraints and horizontal constraints, which include constraints arising from employment contracts
(such as satisfying minimum and maximum thresholds and limiting consecutive assignments). Unlike
our problem, Bilgin et al consider coverage constraints to be soft constraints, in order to prevent
over-constraining of the problem. They incorporate nurse requests into the model formulation,  %SM: tell us who 'they' are :) 
similar to our problem, but they allow requests for both time off and time on.   %rephrase for clarity - not sure I followed?

[...]