### 2021-02-10 paper/tbl/2019_schedule_comparison.tex

\begin{table}[htbp]
  \centering
  \caption{Comparison of the automatically generated (ILP solution) schedule and the final schedule after manual adjustments for 2019.}%
  \label{tbl:2019-schedule-comparison}%
	\begin{adjustbox}{scale=0.8}
	\begin{tabular}{c||ccc||ccc}
		\multicolumn{1}{c||}{\multirow{2}[1]{*}{Week \#}} &                                                                                      \multicolumn{3}{c||}{ILP Solution}                                                                                      &                                                                                    \multicolumn{3}{c}{Manual Adjustment}                                                                                     \\
		                                                  &                                HIV                                 &                                 ID                                 &                              Weekend                               &                                HIV                                 &                                 ID                                 &                              Weekend                               \\ \midrule\midrule
		                        1                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
		                        2                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
		                        3                         &                   \cellcolor[rgb]{ .8,  .2,  0}L                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
		                        4                         &                   \cellcolor[rgb]{ .8,  .2,  0}L                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   \\
		                        5                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
		                        6                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
		                        7                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   \\
		                        8                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   \\
		                        9                         & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
		                       10                         & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .6,  .8,  0}K                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
		                       11                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
		                       12                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
		                       13                         &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
		                       14                         &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
		                       15                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   \\
		                       16                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ 1,  .6,  .8}J                   \\
		                       17                         & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .8,  .2,  0}L                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .8,  .2,  0}L                   \\
		                       18                         & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .8,  .2,  0}L                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
		                       19                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
		                       20                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
		                       21                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
		                       22                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .957,  .69,  .518}E                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
		                       23                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
		                       24                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
		                       25                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
		                       26                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
		                       27                         & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
		                       28                         & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
		                       29                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   \\
		                       30                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                   \cellcolor[rgb]{ 1,  .6,  .8}J                   \\
		                       31                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
		                       32                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
		                       33                         & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .6,  .8,  0}K                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
		                       34                         & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .6,  .8,  0}K                   \\
		                       35                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
		                       36                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
		                       37                         &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
		                       38                         &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
		                       39                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
		                       40                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .957,  .69,  .518}E                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                \cellcolor[rgb]{ .518,  .592,  .69}G                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} \\
		                       41                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   \\
		                       42                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                \cellcolor[rgb]{ .518,  .592,  .69}G                &               \cellcolor[rgb]{ .459,  .443,  .443}H                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
		                       43                         &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &               \cellcolor[rgb]{ .557,  .663,  .859}F                &                \cellcolor[rgb]{ .957,  .69,  .518}E                \\
		                       44                         &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &               \cellcolor[rgb]{ .557,  .663,  .859}F                &               \cellcolor[rgb]{ .557,  .663,  .859}F                \\
		                       45                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                \cellcolor[rgb]{ .518,  .592,  .69}G                \\
		                       46                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                \cellcolor[rgb]{ .518,  .592,  .69}G                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   \\
		                       47                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                \\
		                       48                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .459,  .443,  .443}H                \\
		                       49                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &               \cellcolor[rgb]{ .788,  .788,  .788}D                & \cellcolor[rgb]{ .251,  .251,  .251}\textcolor[rgb]{ 1,  1,  1}{I} &                   \cellcolor[rgb]{ .8,  .2,  0}L                   \\
		                       50                         &               \cellcolor[rgb]{ .788,  .788,  .788}D                &               \cellcolor[rgb]{ .459,  .443,  .443}H                &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &               \cellcolor[rgb]{ .788,  .788,  .788}D                &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &                   \cellcolor[rgb]{ .6,  .8,  0}K                   \\
		                       51                         &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .6,  .8,  0}K                   &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                  \cellcolor[rgb]{ 1,  .851,  .4}C                  &                   \cellcolor[rgb]{ .8,  .2,  0}L                   &               \cellcolor[rgb]{ .459,  .443,  .443}H
	\end{tabular}%
\end{adjustbox}\\[1em]
  \label{tab:addlabel}%
\end{table}%